---
title: "The Geo-scripting course"
author: "Jan Verbesselt and Loïc Dutrieux"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: slidy_presentation

---

## Course concept

- The course is freshly developed based on needs from people who deal with spatial data
- A team of `R` & `Python` experts who use Geo-scripting languages every day helped to develop this course
- Feedback is welcome e.g. via the following [form](http://goo.gl/forms/3iFSNgg3Ne to submit ideas, requests, bugs) 

Key course tips:

  - Avoid copy paste, and make sure you understand all steps
  - Learn how to find help

## Course planning

[Planning of the geo-scripting course](https://docs.google.com/document/d/1yTomAgB9Mny-h0ZEd0mhr0nU5_fD5ysZBAvFYqI5Fn4/edit?usp=sharing)

## Team work & helping each other

- Teams of two (e.g. somebody who followed the `python` course and who didn't follow it)
  - select the team (2 persons) and fill in the [form](http://goo.gl/forms/fzlIlQHlPv)
  - select which team will review your excercise
  
- The [Q&A forum](http://www.geo-informatie.nl/courses/grs50406/question2answer/index.php)
    - Try to obtain the highest score (overall score, best question, best answer)
    - Ask at least one good question
    - Answer at least one question

## Geo-scripting learning objectives

- Know how to find help (on spatial data handling functions)
- Use functions from a library
- Find libraries, functions for geo-data
- Solve scripting problems
    - writing a good question!
    - debugging
- Read, write, and visualise spatial data using a script
- Apply learned concepts for your own work 

## Why geo-scripting?

- Reproducible: avoid clicking and you keep track of what you have done
-	Efficient: you can write a script to do something for you e.g. multiple times e.g. automatically downloading data
-	Build your own tools and functions (e.g. raster filters, MODIS download tool, BFAST package)
-	Enable collaboration: sharing scripts, functions, and packages
-	Good for finding errors i.e. debugging

*examples of advantages of geo-scripting to be added here*
e.g. This course is fully written with scripting languages (i.e. R and MarkDown). What are the advantages?

## What  is a scripting language?

- A scripting language (e.g. R and Python) can interpret and automate the execution of tasks which could alternatively be executed one-by-one by a human operator
- A scripting language is the glue, between different commands, functions, and objectives without the need to compile it for each OS/CPU Architecture
- Different from like C/C++/Fortran since these are languages that need to be compiled first.

## Compiling
- Compiling? 
Passing the source code through a compiler, which translates the `high-level language` instructions into `object code`.

```{r, echo=FALSE, fig.align='center', fig.width=6, strip.white=TRUE}
library(png)
library(grid)
img <- readPNG("figs/load_javac.png")
grid.raster(img)
```


## The main scripting languages for GIS and Remote sensing currently are: 

- R
- Python
- GRASS (grass function are included in QGIS)
- JavaScript for geoweb scripting
- Matlab
- IDL (ENVI)
- ...

## Python versus R

- Python is a general purpose programming language
- R is particularly strong in statistical computing and graphics
- Installing libraries in Python is sometimes challenging
- Syntactic differences can be confusing (especially when combining these) 
- There are many R and Python packages for spatial analyses
- Scripts in both languages can be combined:
  - call R from Python using e.g. RPy2
  - call Python from R (http://rpython.r-forge.r-project.org/)

## R in Python

A very simple example using the Python Rpy2 module:

```{r, engine='python', eval = TRUE}
import rpy2.robjects as robjects
pi = robjects.r['pi']
print(pi[0])
```

## Many programs have support for

- Python (Blender, Sketchup, QGIS, MySQL, PostGIS)
- R (GRASS, QGIS, MySQL, PostGIS)
